Tutorial 3b

### 2025-07-08 manuscript/reviews/ResponseReview1.Rmd

> Stoolmiller, M., & Snyder, J. (2006). Modeling heterogeneity in social interaction processes using multilevel survival analysis. Psychological Methods, 11(2), 164–177. https://doi.org/10.1037/1082-989X.11.2.164

Response: We added these references to the revised ms. on page 6.

> Similarly, it would be helpful to make it clear that this manuscript pertains to estimating single-event occurrences rather than recurring events (which are described in the Lougheed and Stoolmiller tutorials mentioned above)

Response: We added a section in the Supplemental Material (section A in the revised ms.) which explains the types of time-to-event data typically obtained in RT tasks (detection, discrimination, and bistable perception tasks). This section makes it clear that our tutorials pertain only to estimating single-event occurrences. It also visualizes recurrent events and includes references to the Lougheed and Stoolmiller tutorials.

@@

> 6. Line 517: What are model weights? Please explain briefly for the reader,
including how the interpretation differs for loo and waic. 
Also, in the provided
reference (Kruz 2023a), I was only able to find a discussion of dWAIC, rather
than WAIC, so could you please include a definition of or reference for dWAIC in
your context? 
Lastly, when reporting the model weights in the example in the
text, I’d suggest including more than 2 digits so that it is clear if the reported
weights are actually 0 and 1 or if rounding has just made them appear so.

Response: We briefly explain model weights for loo and waic on page 28 of the revised ms.
A discussion of WAIC in Kurz (2023a) can be found in section 4.6.4., so we did not include a reference for dWAIC in the revised ms. Lastly, we print more digits when reporting the model weights.

> 7. Tutorials 3a and 3b: Rather than simply stating that the frequentist models did
not converge and resulted in singular fits, it would be helpful to provide more
discussion. Do you have an idea why the models did not converge? Since this
is a tutorial paper, more explanation is needed. These frequentist examples
don’t add much to the paper in their current form, so I’d either suggest either (a)
providing some suggestions to the reader of alternative approaches or specific
strategies that could be used to define a model that will converge or (b) removing
these examples from the main text and directing the reader to the supplement
instead.

Response: We removed the frequentist examples from the main text and direct the reader to the Tutorials on page 37 of the revised ms.

We mention in Tutorials 3a and 3b that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure.

### 2025-07-08 Tutorial_1a.Rmd
update tutorials with bibliography

---
title: "Tutorial 1a"
author: "Sven Panis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In Tutorial 1a we wrangle raw data from Experiment 1 of Panis and Schmidt (2016), and illustrate how to calculate descriptive statistics (functions of discrete time). We use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.

@@

Read in the raw data file from Panis and Schmidt (2016) - can be skipped.

@@

# 5. Optional: set up life tables and make plots for data aggregated across participants

@@

data_agg_final <- data_agg_nested %>% 
    mutate(censored  = map(data, censor, 600, 40)) %>%
    mutate(ptb_data  = map(censored, ptb)) %>%          
    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
    mutate(condacc   = map(censored, calc_ca)) %>%      
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
    # set haz_yaxis to .7, first_bin_shown to 3, aggregated_data to TRUE, and Nsubj to 6 
    mutate(plot = map2(.x=lifetable_ca, .y=pid, plot_eha, .7, 3, T, 6)) 
  # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
  #            Recommended to set this to 1 when plotting the data of individual subjects.
  # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
  #.           you can set first_bin_shown to another value, such as 3 (i.e., do not show the 
  #            first3 bins in the plots).
  # aggregated_data: For a dataset with data aggregated across individuals, set to TRUE. 
  #            This only controls the title of the plots.
  # Nsubj: The number of individuals before aggregation. Only used to set the title of the plot.
```

@@

>>>>

---
title: "Tutorial 1a"
author: "Sven Panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In Tutorial 1a we wrangle raw data from Experiment 1 of @panisWhatShapingRT2016, and illustrate how to calculate descriptive statistics (functions of discrete time). We use functional programming to set up life tables, and plot the discrete-time hazard, survivor, probability mass, and conditional accuracy functions, when there is one independent variable.

@@

Read in the raw data file from @panisWhatShapingRT2016 - can be skipped. This file can be found on ResearchGate.

@@

# 5. Optional: set up life tables and make one plot for data aggregated across participants

@@

data_agg_final <- data_agg_nested %>% 
    mutate(censored  = map(data, censor, 600, 40)) %>%
    mutate(ptb_data  = map(censored, ptb)) %>%          
    mutate(lifetable = map(ptb_data, setup_lt)) %>%      
    mutate(condacc   = map(censored, calc_ca)) %>%      
    mutate(lifetable_ca = map2(lifetable, condacc, join_lt_ca)) %>%  
    # set haz_yaxis to .7, first_bin_shown to 3, aggregated_data to TRUE, and Nsubj to 6 
    mutate(plot = map2(.x=lifetable_ca, .y=pid, plot_eha, .7, 3, T, 6)) 
  # haz_yaxis: controls the upper limit of the y-axis of the hazard plot.
  #            Recommended to set this to 1 when plotting the data of individual subjects.
  # first_bin_shown: default = 1. In case the first few bins do not contain data for any person, 
  #.           you can set first_bin_shown to another value, such as 3 (i.e., do not show the 
  #            first3 bins in the plots).
  # aggregated_data: When plotting data aggregated across individuals, set to TRUE. 
  #            This only controls the title of the plots.
  # Nsubj: The number of individuals before aggregation. Only used to set the title of the plot.
```

@@

# 6. References.

### 2025-07-08 Tutorial_1b.Rmd
update tutorials with bibliography

---
title: "Tutorial 1b"
author: "Sven Panis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of Panis and Schmidt (2016) when there are TWO independent variables (here: mask type and prime type).

@@

Read in the raw data from Panis and Schmidt (2016).

@@

>>>>

---
title: "Tutorial 1b"
author: "Sven Panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

Tutorial 1b illustrates how to use functional programming to set up life tables, and plot the discrete-time hazard, survivor, and conditional accuracy functions, for the data of Experiment 1 of @panisWhatShapingRT2016 when there are TWO independent variables (here: mask type and prime type).

@@

Read in the raw data from @panisWhatShapingRT2016. This file can be found on ResearchGate.

@@

# 5. References.

### 2025-07-08 Tutorial_2a.Rmd
update tutorials with bibliography

---
title: "Tutorial 2a"
author: "Sven Panis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data (section 3). The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent). We performed some prior predictive checks to determine the prior distributions we will employ (section 8). We compare models using WAIC and LOO (section 4), and interpret parameter estimates for one model (section 5). 
We also plot the logit and cloglog link functions (section 6), and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., hazard) scale (section 7).

@@

The predictive accuracy of a set of models can be compared using WAIC and LOO.

@@

```{r model-weights}
model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "loo") %>% round(digits = 3)

model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "waic") %>% round(digits = 3)
```

@@

To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, Solomon Kurz simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.

@@

Following Gelman et al. (2020), we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs. First, sample the prior distributions using sample_prior="only".

@@

>>>>

---
title: "Tutorial 2a"
author: "Sven Panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data (section 3). The data are taken from the first experiment of @panisWhatShapingRT2016, but using only the no-mask trials (with factor prime = blank, congruent, or incongruent). We performed some prior predictive checks to determine the prior distributions we will employ (section 8). We compare models using WAIC and LOO (section 4), and interpret parameter estimates for one model (section 5). 
We also plot the logit and cloglog link functions (section 6), and illustrate how various prior distributions for an intercept on the logit and cloglog scales look on the original probability (i.e., hazard) scale (section 7).

@@

The predictive accuracy of a set of models can be compared using WAIC and LOO [@mcelreathStatisticalRethinkingBayesian2018; @kurzStatisticalRethinkingSecondEd2023].

@@

```{r model-weights}
model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "loo") %>% round(digits = 3) %>% format(nsmall=2)

model_weights(model_M0i, model_M0d, model_M0c, model_M1i, model_M1d, weights = "waic") %>% round(digits = 3) %>% format(nsmall=2)
```

@@

To gain a sense of what prior logit values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale, @kurzAppliedLongitudinalDataAnalysis2023 simulated a large number of draws from the Uniform(0,1) distribution, converted those draws to the log-odds metric, and fitted a Student's t model.
Here we do the same for prior cloglog values: simulate a large number of draws from U(0,1), convert them to the cloglog metric, and fit a skew-normal model (due to the asymmetry of the cloglog link function), to gain a sense of what prior cloglog values would approximate a uniform distribution on the probability (i.e., discrete-time hazard) scale.

@@

Following @gelmanBayesianWorkflow2020, we perform a prior predictive check to see if the binary observations generated from the specified prior distribution(s) reflects our prior beliefs. First, sample the prior distributions using sample_prior="only".

@@

# 9. References.

Heiss, Andrew (2021). “A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel Bayesian Models.” November 10, 2021. https://doi.org/10.59350/wbn93-edb02.

### 2025-07-08 Tutorial_2b.Rmd
update tutorials 2b, 3a, 3b
update tutorials with bibliography

---
title: "Tutorial 2b"
author: "sven panis"
date: "2024-10-18"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data.

@@

@@

>>>>

---
title: "Tutorial 2b"
author: "sven panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In this Tutorial 2b we fit a Bayesian multilevel model to the conditional accuracy data, using the input file created in Tutorial 1a.

@@

See Heiss (2021).

@@

# 5. References.

Heiss, Andrew (2021). “A Guide to Correctly Calculating Posterior Predictions and Average Marginal Effects with Multilievel Bayesian Models.” November 10, 2021. https://doi.org/10.59350/wbn93-edb02.

### 2025-07-08 Tutorial_3a.Rmd
update Tutorials 3a and 3b with 1 sentence
update tutorials with bibliography

---
title: "Tutorial 3a"
author: "Sven Panis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In this file we build a Frequentist hazard regression model for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

@@

>>>>

---
title: "Tutorial 3a"
author: "Sven Panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In Tutorial 3a we build a Frequentist hazard regression model for the first experiment of @panisWhatShapingRT2016 using only the no-mask trials (prime = blank, congruent, or incongruent). Note that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure [@eager2017mixedeffectsmodelsterrible].

@@

# 5. References

### 2025-07-08 Tutorial_3b.Rmd
update Tutorials 3a and 3b with 1 sentence
update tutorials with bibliography

---
title: "Tutorial 3b"
author: "sven panis"
date: "2024-10-18"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of Panis and Schmidt (2016) using only the no-mask trials (prime = blank, congruent, or incongruent).

>>>>

---
title: "Tutorial 3b"
author: "sven panis"
date: "`r Sys.Date()`"
bibliography: "refs_tutorials.bib"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
editor_options: 
  chunk_output_type: console
---

@@

In this tutorial 3b we fit a Frequentist regression model to conditional accuracies, for the first experiment of @panisWhatShapingRT2016 using only the no-mask trials (prime = blank, congruent, or incongruent). Note that it is common for models to fail to converge in lme4 when they have a reasonably complex random effects structure [@eager2017mixedeffectsmodelsterrible].
